### 2016-01-25 paper.Rnw

\documentclass[12pt,BCOR1mm,DIV18]{scrartcl}
\usepackage{lscape}
\usepackage{microtype}
% -------------------------------------------------------------------------------

\usepackage{cite} % Make references as [1-4], not [1,2,3,4]
\usepackage{url}  % Formatting web addresses  
%\usepackage[utf8]{inputenc} %unicode support
%\usepackage{epsfig}
\usepackage{color,hyperref,comment, booktabs}
\urlstyle{rm}
\newcommand{\cc}{\texttt}

\newcommand{\note}[1]{{\color{red}#1}}
\title{{\Large \bf Analysing time course microarray data using Bioconductor: a case study using yeast2 Affymetrix arrays}}

\author{Colin S. Gillespie$^{1,2, \dagger}$, Guiyuan Lei$^{1,2, \dagger}$, Richard J. Boys$^{1,2}$,\\
Amanda Greenall$^{2,3}$ and Darren J. Wilkinson$^{1,2}$}

<<echo=FALSE>>=
library("knitr")
opts_chunk$set(prompt = FALSE,fig.path='graphics/', fig.align='center',fig.lp="")
#knit_hooks$set()
options(width=75)
knit_hooks$set(
    par.nice = function(before, options, envir) {
        if (before && !is.null(options$par.mfrow)){
            par(mar=c(3,3,2,1),  mgp=c(2,0.4,0), tck=-.01,
                cex.lab=.95,cex.axis=0.9, las=1, mfrow=options$par.mfrow)

        } else if (before) {
            par(mar=c(3,3,2,1),  mgp=c(2,0.4,0), tck=-.01,
                cex.lab=.95,cex.axis=0.9, las=1)
        }
    }, 
    crop=hook_pdfcrop)
read_chunk('ExtractIDs.R')
read_chunk('RemoveProbes.R')
@

\begin{document}
\maketitle \thispagestyle{empty}
\newcommand{\m}{\mbox{}}

\begin{center}
\noindent $^1${\small School of Mathematics \& Statistics, Newcastle University, Newcastle upon Tyne, NE1 7RU, UK.}\\
$^2${\small Centre for Integrated Systems Biology of Ageing and Nutrition (CISBAN), Newcastle University, UK.}\\
$^3${\small Institute for Ageing and Health, Newcastle University, Campus for Ageing and Vitality, Newcastle upon Tyne, NE4 5PL, UK.} \\
$^{\dagger}${\small Both authors contributed equally to the work.}
\end{center}
\begin{abstract}
% Do not use inserted blank lines (ie \\) until main body of text.

\textbf{Background:} Large scale microarray experiments are becoming
increasingly routine, particularly those which track a number of different
cell lines through time. This time-course information provides valuable
insight into the dynamic mechanisms underlying the biological processes being
observed. However, proper statistical analysis of time-course data requires
the use of more sophisticated tools and complex statistical models.

\textbf{Findings:} Using the open source CRAN and Bioconductor repositories
for R, we provide example analysis and protocol which illustrate a variety of
methods that can be used to analyse time-course microarray data. In
particular, we highlight how to construct appropriate contrasts to detect
differentially expressed genes and how to generate plausible pathways from the
data. A maintained version of the R commands can be found at
\url{http://www.mas.ncl.ac.uk/~ncsg3/microarray/}

\textbf{Conclusions:} CRAN and Bioconductor are stable repositories that
provide a wide variety of appropriate statistical tools to analyse time course
microarray data.
\end{abstract}

\section{Introduction}

As experimental costs decrease, large scale microarray experiments are becoming
increasingly routine, particularly those which track a number of different cell
lines through time. This is because time-course information provides valuable
insight into the dynamic mechanisms underlying the biological processes being
observed. However, a proper statistical analysis of time-course data requires
the use of more sophisticated tools and complex statistical models. For example,
problems due to multiple comparisons are increased by catering for changing
effects over time. In this case study, we demonstrate how to analyse time-course
microarray data by investigating a data set on yeast. We discuss issues related
to normalisation, extraction of probesets for specific species, chip
quality,differential expression and network inference. The freely available
software system R (see \cite{R, Gentleman04}) has many benefits for analysing
data of this type and so throughout the analysis we give the R commands that
produce the numerical/graphical output shown in this paper. A maintained version
of the R commands can be found at
\url{http://www.mas.ncl.ac.uk/~ncsg3/microarray/}

\subsection{Description of the data}

The data were collected according to the experimental protocol described in
\cite{Greenall08}. Briefly, three biological replicates were studied on each of
a wild-type (WT) yeast strain and a strain carrying the \textit{cdc13-1}
temperature sensitive mutation (in which telomere uncapping is induced by growth
at temperatures above around 27${^\circ}$C). These replicates were sampled
initially at 23${^\circ}$C (at which \textit{cdc13-1} has essentially WT
telomeres) and then at 1, 2, 3 and 4 hours after a shift to 30${^\circ}$C to
induce telomere uncapping. The thirty resulting RNA samples were hybridised to
Affymetrix yeast2 arrays. The microarray data are available in the ArrayExpress
database (see \cite{Parkinson2009}) under accession number E-MEXP-1551 .

\section{Loading microarray data into Bioconductor}

\subsection{Installing Bioconductor and associated packages}

Assuming that R is already installed, Bioconductor is fairly
straightforward to obtain installation script, viz:

<<eval=FALSE>>=
url = 'http://bioconductor.org/biocLite.R'
source(url)
biocLite()
@

\noindent This installs a number of base packages, including \cc{affy}, \cc{affyPLM},
\cc{limma}, and \cc{gcrma} (see \cite{affy,Bolstad05,Smyth05}). Additional
non-standard packages can also be easily installed. For example, the additional
packages needed for this paper can be installed by using

<<eval=FALSE,tidy=FALSE>>=
##From Bioconductor
biocLite(c('ArrayExpress', 'Mfuzz, 'timecourse', 'yeast2.db', 
    'yeast2probe', 'yeast2cdf', aAffyPLM'))
##From cran
install.packages(c('GeneNet', 'gplots'))
@

\noindent Bioconductor packages are updated regularly on the web and so users
can easily update their currently installed packages by starting a new
R session and then using

<<eval=FALSE>>=
update.packages(repos=biocinstallRepos())
@

\noindent See \cite{Install} for further details on installation. A list of packages
used in this paper is given in Appendix C.

\subsection{Entering data into Bioconductor}

The data used in this paper can be downloaded from ArrayExpress into R using the commands

<<eval=TRUE,hide=TRUE,cache=TRUE,message=FALSE>>=
library(ArrayExpress)
yeast.raw = ArrayExpress('E-MEXP-1551')
@

\noindent A brief description of the \cc{yeast.raw} object can be obtained by using the 
\cc{print(yeast.raw)} command:

{\small\begin{verbatim}
    AffyBatch object
    size of arrays = 496x496 features (3163 kb)
    cdf = Yeast_2 (10928 affyids)
    number of samples = 30
    number of genes = 10928
    annotation = yeast2
\end{verbatim}}
\noindent If the Affymetrix microarray data sets have been downloaded into a single directory, then the \cc{.cel} files can be loaded into R using the \cc{ReadAffy} command.

Also available from ArrayExpress are the experimental conditions. However, some preprocessing is necessary:

<<eval=TRUE,cache=TRUE,tidy=FALSE>>=
ph = yeast.raw@phenoData
exp_fac = data.frame(data_order = 1:30, 
    strain = ph@data$Factor.Value..phenotype.,  
    replicates = ph@data$Characteristics..individual.,
    tps = ph@data$Factor.Value..time.)
levels(exp_fac$strain) = c('m', 'w')
exp_fac = with(exp_fac, exp_fac[order(strain, replicates, tps), ])
exp_fac$replicate = rep(1:3, each=5, 2)
@

\noindent The data frame \cc{exp\_fac} stores all the necessary information, such as strain, time and replicate, which are necessary for the statistical analysis.

Note that there are two yeast species on this chip, \textit{S. pombe}
and \textit{S. cerevisiae}. Also, amongst the 10,928 probesets (with each
probeset having 11 probe pairs), there are 5,900 \textit{S. cerevisiae}
probesets.

\section{Pre-processing}

\subsection{Extraction of S. cerevisiae probesets}

As these microarrays contain probesets for both \textit{S. cerevisiae}
and \textit{S. pombe}, we first need to extract the \textit{S. cerevisiae}
data before normalisation. This can be done by filtering out the \textit{
S. pombe} data using the \verb$s_cerevisiae.msk$ file from the
Affymetrix website (see \cite{AffymetrixA}). Note that users first need to register with the Affymetrix website before downloading this file. Also note that in our analysis, the transcript id i.e. the systematic orf name (obtained from \cite{AffymetrixB}) is used for genes with no name.

We obtain a data frame containing lists of \textit{S. cerevisiae} genes, probes and transcripts (using the function \texttt{ExtractIDs()} in Appendix A) as follows

<<eval=TRUE,message=FALSE,warning=FALSE>>=
#Read in the mask file
s_cer = read.table('s_cerevisiae.msk', skip=2, 
stringsAsFactors=FALSE)
source('ExtractIDs.R')
c_df = ExtractIDs(s_cer[ ,1])
@
\noindent We also need to restrict the view of \cc{yeast.raw} to the $x-$ and $y-$coordinates of the \textit{S. cerevisiae} probesets in the \cc{cdf} environment by using
<<eval=TRUE,message=FALSE>>=
#Get the raw dataset for S. cerevisiae only
library(affy)
library(yeast2probe)
source('RemoveProbes.R')
cleancdf = cleancdfname(yeast.raw@cdfName)
RemoveProbes(s_cer[, 1], cleancdf, 'yeast2probe') 
@
Note that the commands in \cc{RemoveProbes.R} are listed in Appendix A. Thus the attributes of \cc{yeast.raw}, obtained via
\cc{print(yeast.raw)}, are now
{\small\begin{verbatim}
    AffyBatch object
    size of arrays = 496x496 features (3167 kb)
    cdf = Yeast_2 (5900 affyids)
    number of samples = 30
    number of genes = 5900
    annotation = yeast2
\end{verbatim}}
\noindent and the number of genes (actually probesets here) is 5,900 now that the \textit{S. pombe} probesets have been removed.

\subsection{Data Quality Assessment}

Before any formal statistical analysis, it is important to check for
data quality. Initially, we might examine the perfect and mismatch
probe-level data to detect anomalies. Images of the first five arrays can be obtained using
<<s_fig1,eval=FALSE,tidy=TRUE, echo=-1>>=
par(mar=c(3,3,2,1), mgp=c(2,0.4,0), tck=-.01, cex.axis=0.9, las=1, mfrow=c(2,3))
for(i in 1:5) {
    plot_title = paste('Strain: ', exp_fac$strain[i], 
        'Time: ', exp_fac$tps[i])
        d = exp_fac$data_order[i]
    image(yeast.raw[,d], main=plot_title) 
}
@

\noindent These commands produce the image shown in Appendix B: Figure \ref{F1A}. Data
quality can be assessed by examining such images for anything that appears
non-random such as rings, shadows, lines and strong variations in shade. The
images for our data set do not appear to have any non-random structure and so
data quality is probably high.

Another useful quality assessment tool is to examine density plots of
the probe intensities. The command

<<s_fig2,eval=FALSE,tidy=FALSE,echo=-1>>=
par(mar=c(3,3,2,1), mgp=c(2,0.4,0), tck=-.01, cex.axis=0.9, las=1)
d = exp_fac$data_order[1:5]
hist(yeast.raw[,d], lwd=2, ylab="Density", xlab="Log (base 2) intensities") 
@
\noindent produces the image shown in Appendix B: Figure \ref{F2A}. Typically,
differences in spread and position are corrected by normalisation. However, the
appearance of significant multi-modality in the distribution or many outlying
observations are indicative of poor data quality.

Other exploratory data analysis techniques that should be carried
include MAplots, where two microarrays are compared and their log
intensity difference for each probe on each gene are plotted against
their average. Also of interest is to examine RNA degradation (see
\cite{Bolstad05}), although \cite{Archer06} cast some doubt over the
validity of this method. For details on how to carry out both of these
methods in R, see \cite{Alvord07, speedcourse} for detailed
instructions.

\subsection{Normalising Microarray Data}

There are number of methods for normalising microarray data. Two of the most
popular methods are GeneChip RMA (GCRMA) and Robust Multiple-array Average
(RMA); see \cite{Irizarry03,Wu04}. Essentially, GCRMA and RMA differ in how they
deal with background noise, with GCRMA using a more sophisticated correction
algorithm. However, the approach adopted by GCRMA means that it can be
time-consuming to use with large data sets in contrast to RMA. A potential
drawback of using RMA is that it assumes that the overall levels of expression
are similar for each array. However this assumption may be invalid if, for
example, mutant cells have a radically different level of transcriptional
activity than the WT. For further information regarding normalising microarray
data sets, see for example \cite{Harr06, Labbe07}.

Since we have thirty microarray data sets and believe that the levels of transcriptional activity are similar across strains, we will use the RMA normalisation method. This technique normalises
across the set of hybridizations at the probe level. The data can be
normalised via

<<eval=TRUE,cache=TRUE,results=FALSE,message=FALSE>>=
yeast.rma = rma(yeast.raw)
yeast.matrix = exprs(yeast.rma)[, exp_fac$data_order]
colnames(yeast.matrix) = paste0(exp_fac$strain, exp_fac$tps)
exp_fac$data_order = 1:30
@

The normalisation procedure consists of three steps: model-based
background correction, quantile nomralisation and robust
averaging. The aim of the quantile normalisation is to make the
distribution of probe intensities for each array in a set of arrays
the same. We illustrate its effect by studying boxplots of the raw
\textit{S. cerevisiae} data against their normalised counterparts
values, shown in the Appendix B: Figure \ref{F3A}. Boxplots provide a useful graphical
view of data distributions and contain their median, quartiles,
maximum and minimum values. The \cc{boxplot} command is in the
\cc{affyPLM} package and so the figure is produced by using

<<s_fig3,eval=FALSE, echo=-1 >>=
par(mar=c(3,3,2,1), mgp=c(2,0.4,0), tck=-.01,cex.axis=0.9, las=1, mfrow=c(1,2))
library(affyPLM)
#Raw data intensities
boxplot(yeast.raw, col='red', main='', ylim=c(2,16))
#Normalised intensities
boxplot(yeast.rma, col='blue', ylim=c(2,16))
@

\subsection{Principal Component Analysis}

Principal component analysis (PCA) is useful in exploratory data
analysis as it can reduce the number of variables to consider whilst
still retaining much of the variability in the data. In particular,
PCA is useful for identifying patterns in the data. Essentially,
principal components partition the data into orthogonal linear
components which explain different contributions to the variability in
the data. The first component explains the largest contribution to
variability in the original dataset, that is, retains most
information, with the second component explaining the next largest
contribution to variability, and so on. The following commands calculate the principal components

<<eval=TRUE>>=
yeast.PC = prcomp(t(yeast.matrix))
yeast.scores = predict(yeast.PC)
@

\noindent which we can then plot using

<<F1,eval=TRUE, tidy=FALSE,dev="pdf",fig.pos="!t", cache=TRUE, fig.width=6.5,fig.height=4,out.width="0.8\\textwidth",fig.cap="A plot of the first two principal components. The red symbols correspond to the wild-type strain.", par.nice=TRUE>>=
#Plot of the first two principal components
plot(yeast.scores[,1], yeast.scores[,2], xlab='PC 1', ylab='PC 2', 
    pch=rep(1:5, 6), col=as.numeric(exp_fac$strain))
legend("bottomleft", pch=1:5, cex=0.6, 
    c('t 0', 't 60', 't 120', 't 180', 't 240'))
@

Figure \ref{F1} highlights a clear (and expected) time effect in the mutant
yeast which is not present in the wild-type strain. In particular,
mutant samples are clustered by their time points; for example, the
three mutant replicates at time point~4 are clustered at the bottom
right of the figure.

\section{Identifying differentially expressed genes}

In this experiment, interest lies in differences in gene expression
over time between the wild-type and mutant yeast strains. It is
expected that the wild-type expression level is independent of
time. Also we anticipate that the mutant expressions at time $t=0$ are
the same as the wild-type expression level. This hypothesis is
supported by the PCA plot in Figure~\ref{F1}.

There are currently two main packages available to detect
differentially expressed genes using this kind of data: the
\texttt{timecourse} package and the \texttt{limma} package. We illustrate
how to analyse these data using both packages.

\subsection{Using the timecourse package}

This package assesses treatment differences by comparing time-course
mean profiles allowing for variability both within and between time
points. It uses the multivariate empirical Bayes model proposed by
\cite{Tai06}. Further details of the \texttt{timecourse} package can be
found in \cite{Tai07}. After installing the \texttt{timecourse} package,
we construct a \texttt{size} matrix describing the replication
structure using
<<eval=TRUE,message=FALSE>>=
library(timecourse)
size = matrix(3, nrow = 5900, ncol = 2)
@
\noindent To extract a list of differentially expressed we calculate the
Hotelling statistic $\tilde T^2$ via
<<eval=TRUE,tidy=FALSE, cache=TRUE>>=
c.grp = as.character(exp_fac$strain)
t.grp = as.numeric(exp_fac$tps)
r.grp = as.character(exp_fac$replicate)
MB.2D = mb.long(yeast.matrix, times = 5, method = '2', 
    reps = size, condition.grp = c.grp, time.grp = t.grp, 
    rep.grp = r.grp)
@
\noindent The top (say) one hundred genes can be extracted via
<<eval=TRUE,cache=TRUE>>=
gene_positions = MB.2D$pos.HotellingT2[1:100]
gnames = rownames(yeast.matrix)
gene_probes = gnames[gene_positions]
@
\noindent The expression profiles can also be easily obtained. The profile for the top ranked expression is found using
<<s_fig4,eval=FALSE, echo=-1>>=
par(mar=c(3,3,2,1), mgp=c(2,0.4,0), tck=-.01, cex.axis=0.9, las=1)
plotProfile(MB.2D, ranking=1, gnames=rownames(yeast.matrix))
@

\noindent and is shown in the Appendix B: Figure \ref{F4A}.

\subsection{Using the limma package}

The \cc{limma} package uses the moderated $t$-statistic described by
\cite{Smyth04,Smyth05}. The function \cc{lmFit} within the \cc{limma}
library fits a linear model for each gene for a given series of
arrays, where the coefficients of the fitted models describe the
differences between the RNA sources hybridised to the
arrays. Precisely, we fit the model $E[y_g] = X \alpha_g$, where
$y_g=(y_{g,1}, \ldots, y_{g,n})^T$ contains the expression values for
gene~$g$ across the $n$ arrays, $X$ is a design matrix which describes
key features of the experimental design used and $\alpha_g$ is the
coefficient vector for gene~$g$. In the analysis studied here, the
yeast data consists of data from $n=30$ arrays. The entries in the
columns of $X$ depend on the experimental design used: there are two
yeast strains (mutant and wild type), each measured at five separate
time points, and we are interested in comparing the gene expressions
between mutant and wild type strains over time. Thus we seek a linear
model describing the ten strain$\times$time combinations by
determining values for the ten coefficients in the coefficient vector
$\alpha_g$. We will label these ten coefficients as (`m0', `m60, `m120',
`m180', `m240', `w0', `w60', `w120', `w180', `w240'), where the first five
coefficients represent the levels of the mutant strain at time points
$t=0,1,2,3,4$ and the remaining five coefficients are the
equivalent versions for the wild type strain. Statistically speaking,
the model has a single factor with ten levels. The design matrix $X$
links these factors to the data in the arrays by having zero entries
except when an array contributes an observation to a particular
strain$\times$time combination. For example, array~$26$ measures the
expression of the first wild type microarray at time $t=0$ and so
contributes an observation to level `w0', the sixth strain$\times$time
combination. Thus the entry in row~26, column~6 of the design matrix
$X(26,6)=1$. Further, the arrays are arranged in groups of three
replicates. Thus the overall experimental structure
(\verb$expt_structure$ below) has three arrays on level `m0', then
three arrays on `m60', and so on.  Setting up the factor levels and the
design matrix is done in R by using

<<eval=TRUE,tidy=FALSE, message=FALSE>>=
library(limma)
expt_structure = factor(colnames(yeast.matrix))

#Construct the design matrix
X = model.matrix(~0 + expt_structure)
colnames(X) =  c('m0', 'm60', 'm120', 'm180', 'm240', 
    'w0', 'w60', 'w120', 'w180', 'w240')
@

\noindent and then the coefficient vector $\alpha_g$ is estimated via the command

<<eval=TRUE,cache=TRUE>>=
lm.fit = lmFit(yeast.matrix, X) 
@

Determining the differentially expressed genes amounts to studying
contrasts of the various strain$\times$time levels, as described by
a contrast matrix $C$. For these data, we are mainly interested in
differences at the later time points, and so a possible set of
contrasts to investigate is that of differences between the mutant and
wild type strains at each time point, that is, \texttt{('m60-w60', 'm120-w120', 'm180-w180', 'm240-w240')}. The \verb"limma" package allows complete flexibility over the choice of contrasts, however this necessarily includes an additional level of complexity. The values in the coefficient
vector of contrasts, $\beta_g=C^T \alpha_g$ for gene~$g$, describe the
size of the difference between strains at each time point. The
relevant R commands are

<<eval=TRUE,cache=TRUE>>=
mc = makeContrasts('m60-w60', 'm120-w120', 'm180-w180', 'm240-w240', levels=X)
c.fit = contrasts.fit(lm.fit, mc)
eb = eBayes(c.fit)
@

The final command uses the \cc{eBayes} function to produce moderated
$t$-statistics which assess whether individual contrast values
$\beta_{gj}$ are plausibly zero, corresponding to no significant evidence of a
difference between strains at time point~$j$. The moderated $t$-statistic is constructed using a shrinkage approach and so is not as sensitive as the standard $t$-statistic to small sample sizes. It also gives a
moderated $F$-statistic which can be used to test whether all
contrasts are zero simultaneously, that is, whether there is no
difference between strains at all time points.

\subsection{Ranking differentially expressed genes}

There are a number of ways to rank the differentially expressed
genes. For example, they can be ranked according to their log-fold
change

<<eval=TRUE,results='hide',cache=TRUE>>=
#see help(toptable) for more options
toptable(eb, sort.by='logFC') 
@

\noindent or by using $F$-statistics

<<eval=TRUE, results='hide',cache=TRUE>>=
topTableF(eb)
@

\noindent The advantage of using $F$-statistics over the log fold change is that the $F$-statistic takes into account variability and reproducibility, in addition to fold-change.

Our analysis is based on a large number of statistical tests, and so we must correct for this multiple testing. In our example we use the (very) conservative Bonferroni correction since we have a large number of differentially expressed genes and the resulting corrected list is still long. Another common method of correcting for multiple testing is to use the false discovery rate (fdr) (use the command \texttt{?p.adjust} to obtain further details). The following commands rank genes according to
their (corrected) $F$-statistic $p$-value and annotates the output by
indicating the direction of the change for each contrast for each
gene: $+1$ for up-regulated expression (mutant type having higher
expression than wild type at a particular time point), $-1$ for
down-regulated expression and $0$ for no significant change.

<<eval=TRUE,cache=TRUE,tidy=FALSE>>=
modFpvalue = eb$F.p.value
##Change 'bonferroni' to 'fdr' to use the 
##false discovery rate as a cut-off
indx = p.adjust(modFpvalue, method='bonferroni') < 0.05
sig = modFpvalue[indx]

#No. of sig. differential expressed genes
nsiggenes = length(sig) 
results = decideTests(eb, method='nestedF')

modF = eb$F
modFordered = order(modF, decreasing = TRUE)

#Retrieve the significant probes and genes
c_rank_probe = c_df$probe[modFordered[1:nsiggenes]]
c_rank_genename = c_df$genename[modFordered[1:nsiggenes]]

#Create a list and write to a file
updown = results[modFordered[1:nsiggenes],]
write.table(cbind(c_rank_probe, c_rank_genename, updown), 
    file='updown.csv', sep=',', 
    row.names=FALSE, col.names=FALSE)
@

The following code (adapted from lecture material found at
\cite{speedcourse}) plots the time course expression for the top
one hundred differentially expressed genes according to their
$F$-statistic (see Figure~\ref{F2})

<<F2,eval=TRUE,echo=FALSE,dev="pdf",fig.pos="!t", cache=TRUE, fig.width=6.5,fig.height=5,out.width="0.8\\textwidth",fig.cap="Time course expression levels for the top 9 differentially expressed genes, ranked by their $F$-statistic. The triangles and circles correspond to the wild-type and mutant genes respectively.",par.nice=TRUE, par.mfrow=c(3,3), tidy=FALSE>>=

for(i in 0:8){
    indx = rank(modF) == nrow(yeast.matrix)-i

    id = c_df$probe[indx]
    name = c_df$genename[indx]
    exprs.row = yeast.matrix[indx,]
    genetitle = paste(sprintf('%.30s', id), sprintf('%.30s', name), 
                      'Rank =', i+1)

    plot(0, pch=NA, xlim=range(0, 240), ylim=range(exprs.row), 
         ylab='Expression', xlab='Time', main=genetitle,
         cex.main=0.7)

    for(j in 1:6){
        pch_value = as.character(exp_fac$strain[5*j])
        points(c(0, 60, 120, 180, 240), exprs.row[(5*j-4):(5*j)], 
               type='b', pch=pch_value)
    }  
} 
@

<<eval=FALSE, tidy=FALSE>>=
#Rank of Probesets, also output gene names
par(mfrow=c(3, 3), ask=TRUE)
for(i in 0:99){
    indx = rank(modF) == nrow(yeast.matrix) - i
    id = c_df$probe[indx]
    name = c_df$genename[indx]
    exprs.row = yeast.matrix[indx,]
    genetitle = paste(sprintf('%.30s', id), sprintf('%.30s', name), 
                      'Rank =', i+1)

    plot(0, pch=NA, xlim=range(0, 240), ylim=range(exprs.row), 
         ylab='Expression', xlab='Time', main=genetitle)

    for(j in 1:6){
        pch_value = as.character(exp_fac$strain[5*j])
        points(c(0, 60, 120, 180, 240), exprs.row[(5*j-4):(5*j)], 
               type='b', pch=pch_value)
    }  
} 
@

When interpreting rank orderings based on statistical significance, it
is important to bear in mind that a statistically significant
differential expression is not always biologically meaningful. For
example, Figure~\ref{F2} contains \textit{RNR2}. This gene is highly
significant because of low variation in its time course. However the
actual difference in expression levels between wild-type and mutant
stains is relatively small. We address this problem in the next
section.

\subsection*{Comparison of the \cc{timecourse} and \cc{limma} packages}

Both packages have different strengths. One advantage of
the \verb"timecourse" package over the \verb"limma" package is that it
allows for correlation between repeated measurements on the same experimental unit, thereby reducing false
positives and false negatives; these false positives/negatives are a
significant problem when the variance-covariance matrix is poorly
estimated. An advantage of the \cc{limma} package is that it allows
more flexibility by allowing users to construct different
contrasts. In general we might expect both packages to produce fairly
similar lists of say the top 100 probesets. In the analysis of the
yeast data, we can determine the overlap of the top 100 probesets by
using

<<eval=TRUE, cache=TRUE>>=
N = 100
gene_positions = MB.2D$pos.HotellingT2[1:N]
tc_top_probes = gnames[gene_positions]
lm_top_probes = c_df$probe[modFordered[1:N]]
length(intersect(tc_top_probes, lm_top_probes))
@

\noindent The result is a moderately large overlap of fifty-three probesets. We
note that changing the ranking method in the \texttt{limma} package also
yields similar results as those given by the
\texttt{timecourse} package.

\subsection{Two fold-change list}

When looking for ``interesting'' genes it can be helpful to restrict
attention to those differential expressed that are both statistically
significant and of biological interest. This objective can be achieved
by considering only significant genes which show, say, at least a
two-fold change in their expression level. This gene list is obtained using the following code (adapted from \cite{Alvord07})

<<eval=TRUE,cache=TRUE>>=
#Obtain the maximum fold change but keep the sign 
maxfoldchange = function(foldchange){
    foldchange[which.max(abs(foldchange))]
}

difference = apply(eb$coeff, 1, maxfoldchange)
pvalue = eb$F.p.value
lodd = -log10(pvalue)

#hfc: high fold-change
nd = (abs(difference)>log(2, 2))
ordered_hfc = order(abs(difference), decreasing=TRUE)
hfc = ordered_hfc[1:length(difference[nd])]

np = p.adjust(pvalue, method='bonferroni') < 0.05

#lpv: low p-value (large F-value)
ordered_lpv = order(abs(pvalue), decreasing=FALSE)
lpv = ordered_lpv[1:length(pvalue[np])]

oo = union(lpv, hfc) 
ii = intersect(lpv, hfc) 
@

Figure~\ref{F3} contains a ``volcano'' plot which illustrates the effect of using different
levels of fold change and significance thresholds. The figure is 
produced by using the following code

<<F3,eval=TRUE,dev="pdf",fig.pos="!t", cache=TRUE, fig.width=6.5,fig.height=4,out.width="0.8\\textwidth",fig.cap="Volcano plot showing the bonferroni cut-off and the two-fold change.", nice.par=TRUE,tidy=FALSE>>=
#Construct a volcano plot using moderated F-statistics
plot(difference[-oo], lodd[-oo], xlim=range(difference), 
ylim=range(lodd), cex=0.7)
points(difference[hfc], lodd[hfc], pch=18, cex=0.7) 
points(difference[lpv], lodd[lpv], pch=1, cex=0.7)

#Add the cut-off lines
abline(v=log(2, 2), col=5); abline(v=-log(2, 2), col=5)
abline(h=-log10(0.05/5900), col=5)

text(min(difference) + 1, -log10(0.05/5900) + 0.2, 
    'Bonferroni cut off', cex=0.8) 
text(1, max(lodd) - 1, paste(length(ii), 'intersects'), cex=0.8)
@

\section{Cluster Analysis}

Biological insight can be gained by determining groups of
differentially expressed genes, that is, groups of genes which
increase or decrease simultaneously. This can be achieved by using
cluster analysis.

\subsection{Traditional cluster analysis}

In this section, we separate the top fifty differentially expressed
genes into groups of similar pattern (clusters). Clearly different
genes will have different overall levels of expression and so we first
standardise their measurements by taking the expression level of the
mutant strain (at each time point) relative to the wild-type at time
$t=0$:

<<eval=TRUE>>=
c_probe_data = yeast.matrix[ii,]
#Average of WT
wt_means = apply(c_probe_data[,16:30], 1, mean)
m = matrix(nrow=dim(c_probe_data)[1], ncol=5)

for(i in 1:5){
    mut_rep = c(i, i+5, i+10)
    m[ ,i] = rowMeans(c_probe_data[ ,mut_rep]) - wt_means
}
colnames(m) = sort(unique(exp_fac$tps))
@

\noindent The heatmap in Figure~\ref{F4} is
obtained by using the function \cc{heatmap.2} from the library
\cc{gplots} via the following code

<<F4,eval=TRUE,dev="pdf",fig.pos="!t", cache=TRUE, fig.width=6.5,fig.height=4,out.width="0.8\\textwidth",fig.cap="Clustering of the top fifty differentially expressed genes. Red and green correspond to up- and down-regulation respectively.",message=FALSE, nice.par=TRUE,tidy=FALSE>>=
library(gplots)
#Cluster the top 50 genes
heatmap.2(m[1:50,], dendrogram ='row', Colv=FALSE, col=greenred(75), 
          key=FALSE, keysize=1.0, symkey=FALSE, density.info='none', 
          trace='none', colsep=rep(1:10), sepcolor='white', 
          sepwidth=0.05, labRow = NA, cexCol=1,
          hclustfun=function(c){hclust(c, method='average')})
@

Figure \ref{F4} shows the relative expression levels for the mutant strain
at each time point (`0', `60', `120', `180', `240'). As expected, the
relative expression levels at time $t=0$ are very similar. However, as
time progresses, groupings of genes appear whose levels are
up-regulated (red) or down-regulated (green). Note that the intensity
of the colour corresponds to the magnitude of the relative expression. Gene
names appear on the right side of the figure and on the left side, the
cluster dendrogram shows which genes have similar expression. The
dendrogram suggests that there are perhaps six to ten clusters.

\subsection{Soft clustering}

Soft clustering methods have the advantage that a probe can be
assigned to more than one cluster. Furthermore, it is possible to
grade cluster membership within particular groupings. Soft clustering
is considered more robust when dealing with noisy data; for more details
see \cite{Futschik05,Futschik07}. The \cc{Mfuzz} package
implements soft clustering using a fuzzy c-means algorithm. Analysing
the data for $c=8$ clusters is achieved by using

<<F5,eval=TRUE,dev="pdf",fig.ext='pdf',fig.pos="!t", cache=TRUE, fig.width=6.5,fig.height=4,out.width="0.8\\textwidth",fig.cap="Eight clusters obtained from the \\texttt{Mfuzz} package.",message=FALSE>>=
library(Mfuzz)
tmp_expr = new('ExpressionSet', exprs=m)
cl = mfuzz(tmp_expr, c=8, m=1.25)
mfuzz.plot(tmp_expr,cl=cl, mfrow=c(2, 4), new.window = FALSE)
@
Of course, it is usually not clear how many clusters there are (or
should be) within a dataset and so the sensitivity of conclusions to
the choice of number of clusters ($c$) should always be
investigated. For example, if $c$ is chosen to be too large then some
clusters will appear sparse and this might suggest choosing a smaller
value of~$c$. Figure~\ref{F5} shows the profiles of the
eight clusters obtained from the \cc{Mfuzz} package. The probes 
present within each cluster can be found by using

<<eval=TRUE,results='hide'>>=
cluster = 1
cl[[4]][, cluster] 
@

\section{Genetic regulatory network inference}

Recent research in the analysis of microarray experiments has produced
several methods for determining plausible transcriptional regulation networks from the data; see,
for example, \cite{Dobra04,Husmeier03,Yu04,Lee04}. Such networks can
provide valuable insight into the underlying biological mechanisms
producing the data. The networks typically consist of nodes
(representing genes or proteins) and edges between nodes (representing
relationships between genes). Methods also exist for inferring dynamic
gene association networks in which the direction of causation is
represented by an arrow. One such method (with an easy-to-use R
package developed by the Strimmer lab) uses a shrinkage approach to
calculate the partial correlation coefficient and works for both
static and dynamic (time-course) data; see \cite{Strimmer05,
Strimmer06, Strimmer07}. Their heuristic algorithm is fast and so
provides a quick insight in the structure of the network. It works by
first creating a \cc{longitudinal} R object. In our illustration, we
use data on the top one hundred differentially expressed genes. The
data are stored in a matrix \cc{m}, where the rows are genes and the
fifthteen columns are the arrays. In order to use the R functions in
their package, as we have time course data rearrange the row order according to the time points. The first three
rows of the resulting matrix \cc{mnew} are data on the three mutant arrays at
time $t=0$, the next three arrays at time point $t=60$, and so on. This
is achieved by using the commands

<<eval=TRUE,results='hide',message=FALSE>>=
exp_fac = with(exp_fac, exp_fac[order(strain, tps, replicates), ])
#Construct a longitudinal object
library(GeneNet)
ngenes = 100
m = yeast.matrix[ii[1:ngenes],]
mnew = m[,exp_fac$data_order[1:15]]
mlong = as.longitudinal(t(mnew), repeats=3, time=0:4)
@

\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{graphics/network.jpg}
\caption{Gene network inferred from the yeast microarray data
set. Black and grey indicate positive and negative (partial)
correlation respectively.}
\label{F6}
\end{figure}
\noindent Next the partial correlations are computed and then (local) values are
assigned to all possible edges, from which the important edges can be
determined according to a threshold criteria. Finally a
Graphviz\footnote{See \url{http://www.graphviz.org/} for details.} file is
outputted to generate the graph. Figure~\ref{F6} shows the resulting
network for these data and is generated by using

<<eval=FALSE,cache=TRUE,message=FALSE,results='hide'>>=
#Compute partial correlations
pcor.dyn = ggm.estimate.pcor(mlong, method = 'dynamic')

#Assign (local) fdr values to all possible edges
m.edges = network.test.edges(pcor.dyn, direct=TRUE)

#Construct graph containing top edges
m.net = extract.network(m.edges, method.ggm='number', 
cutoff.ggm=100)

#Construct a Graphviz dot file
rnames = vector('list', length(1))
rnames = c_df$genename[ii[1:ngenes]]
network.make.dot(filename='net.dot', m.net, rnames, 
    main='Yeast Network')
@

\noindent Note that, in the function \cc{ggm.estimate.pcor}, the default method
\cc{static} employs the function \cc{pcor.shrink}, whereas the
\cc{dynamic} method uses \cc{dyn.pcor}. The difference between the two
estimators is that the latter takes the spacings between time points
into account if the input consists of multiple time course data (which
must be provided as a \cc{longitudinal} object).

\section{Conclusion}

The response to telomere uncapping in \textit{cdc13-1} strains was expected to
share features in common with responses to cell cycle progression,
environmental stress, DNA damage and other types of telomere
damage. The statistical analysis determined lists of probesets
associated with genes involved in all of these processes. The
techniques used focussed on making best use of the temporal
information in time-course data.  The use of \textit{cdc13-1} strains, which
uncap telomeres quickly and synchronously, also allowed the
identification of genes involved in the acute response to telomere
damage. This case study has demonstrated the power of R/Bioconductor
to analyse time-course microarray data. Whilst the statistical
analysis of such data is still an active research area, this paper
presents some of the cutting-edge tools that are available to the life
science community.  All software discussed in this article is free,
with many of the packages being open-source and subject to on-going
development.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Competing interests}

The authors declare that they have no competing interests.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Authors contributions}

AG conducted the microarray experiments. All authors participated
in the analysis of the data and in the writing of the manuscript. CSG maintains this document.

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}

We wish to thank Dan Swan (Newcastle University Bioinformatics Support Unit) and David Lydall for helpful
discussions. The authors are affiliated with the Centre for Integrated
Systems Biology of Ageing and Nutrition (CISBAN) at Newcastle
University, which is supported jointly by the Biotechnology and
Biological Sciences Research Council (BBSRC) and the Engineering and
Physical Sciences Research Council (EPSRC).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\appendix

\section{R commands for extracting S. cerevisiae ids, removing unwanted probesets and converting probesets to genes}

\medskip

\subsection*{R function for extracting S. cerevisiae ids}

As these microarrays contain probesets for both \textit{S. cerevisiae}
and \textit{S. pombe}, we first need to extract the \textit{S. cerevisiae}
data before normalisation. This can be done by filtering out the \textit{
S. pombe} data using the \texttt{s\_cerevisiae.msk} file from the
Affymetrix website (see \cite{AffymetrixA}). Note that the transcript
id, i.e. the systematic orf name (obtained from \cite{AffymetrixB}) will be used for genes with no
name.

<<ExtractIDs,tidy=FALSE>>=
@

\clearpage

\subsection*{R function for removing unwanted probesets}

If an Affymetrix microarray chip contains more than one species then
it can be useful to focus on a particular species and filter out the
unwanted probesets. For example, the \cc{yeast2.db} Affymetrix chip
contains both \textit{S. pombe} and \textit{S. cerevisiae} yeast
species. To filter out the \textit{S. pombe} probesets, we remove the
mappings from the $x-$, $y-$coordinates to the \textit{S. pombe}
probesets in the cdf environment. The following function
(adapted from \cite{Alvord07}) removes the unwanted \textit{S. pombe}
instances:

<<RemoveProbes,tidy=FALSE>>=
@

\clearpage

\subsection*{Probeset level to gene level}

It is worth noting that RMA yields probeset expression
levels. However, there can be several probesets that map to a single
gene in an Affymetrix array. If gene level expression is required
(instead of probeset expression) the following code will average over
probesets for each gene

<<eval=FALSE>>=
#Function to average the expression of 
#probesets which map to same gene
probeset2genelevel = function(onesample)
    return(tapply(onesample, factor(c_df$genename), mean))

#Average for each column/array
c_gene_data = apply(exprs(yeast.rma), 2, probeset2genelevel)
@

For the \textit{S. cerevisiae} data, the majority of genes only have a
single probeset. However, thirty five genes have two or three
associated probesets.  In this case study, we will study genes at the
probeset level.
\clearpage

\section{Additional figures}

<<s_fig11,ref.label="s_fig1",eval=TRUE,dev="png",fig.ext="png",cache=FALSE,dpi=300,out.width="0.8\\linewidth",echo=FALSE, fig.cap="Image plots of the mismatch and perfect match probe intensities for the first replication of the mutant yeast strain. The corresponding times are indicated in the plot.", fig.pos="h">>=
#knitr::run_chunk("s_fig1")
@

 \bigskip

<<F2A,ref.label="s_fig2",eval=TRUE,dev="pdf",fig.ext="pdf",cache=TRUE,out.width="0.6\\linewidth",echo=FALSE, fig.cap="Density plots for the first replication of the mutant yeast strain.",nice.par=TRUE,message=FALSE>>=
@

<<F3A,ref.label="s_fig3",eval=TRUE,dev="pdf",fig.ext="pdf",cache=TRUE,out.width="0.8\\linewidth",echo=FALSE, message=FALSE,fig.cap="Boxplots of the raw and normalised intensities. The default boxplot is to include both PM and MM intensities, whereas for the density plots in Figure~\ref{F2} the default is for only the PM intensities.">>=
@

\bigskip
<<F4A,ref.label="s_fig4", eval=TRUE,dev="pdf",fig.ext="pdf",cache=TRUE,out.width="0.6\\linewidth",echo=FALSE, fig.cap="Time course expression levels for the top differentially expressed gene, ranked by their Hotelling statistic using the \texttt{timecourse} library.">>=
@

\clearpage

\section{Session Information}

<<eval=TRUE>>=
sessionInfo()
@

\clearpage

%{\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
\bibliographystyle{plain}  % Style BST file
\bibliography{references}      % Bibliography file (usually '*.bib' )

%%%%%%%%%%%

% Not possible to call purl within a knitr document.
% <<echo=FALSE>>=
% library(knitr)
% purl("paper.Rnw")
% @

\end{document}